Demo arXiv template

### 2022-11-21 paper/paper.tex
first attempt to write notation into pipeline steps

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{article}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
  \setmathfont[]{Latin Modern Math}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{arxiv}
\usepackage{orcidlink}
\usepackage{amsmath}
\usepackage[T1]{fontenc}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
indexes \sep data pipeline \sep 
software design

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, frame hidden, enhanced, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

The construction of natural hazard indexes also follows a set of steps,
which is usually illustrated using a flowchart in the paper.

\begin{itemize}
\tightlist
\item
  constructing time series index should also be encapsulated in the
  framework
\end{itemize}

Here we assume a concept of interest is determined, relevant variables/
indicators are identified and available to construct indexes.

Before the mathmetical construction of indexes

@@

\caption{Diagram of pipeline steps for index construction.}

@@

\emph{Another section on original data directly downloaded, can have
different spatial resolution, temporal granularity, data quality
problem. After processing them and align them together they become the
``raw data''}

The data used to construct the natural hazard index usually have three
dimensions, one for location, one for time, and one for multivariate.
Mathematically, it can be written as \(X_{j, s, t}\), where
\(j = 1, 2, \cdots, J\) for variable, \(s = 1, 2, \cdots, S\) for
location, and \(t = 1, 2, \cdots, T\) for time.

The location \(s\) can refer to vector points or areas characterised by
longitude-latitude coordinates, or raster cells obtained from satellite
images.

The time dimension \(t\) can be daily, weekly, biweekly (14-16 days),
monthly, or even quarterly

Variables

This multidimensional array structure is commonly used in geospatial
analysis

Given the variety of data sources at different spatial resolution and
temporal granularity, the raw data may first come in multiple pieces.
Sometimes, even a considerable amount of work is needed to align the
spatial and temporal extent of multivariate data.

A notation for different variables have different spatial and temporal
granularity \(X_{j_1, s_1, t_1}\)???

@@

mostly happen with raster data

@@

A specific transformation on the scale of the data

z-score standardising, min-max standardisation into {[}0, 1{]} or {[}0,
100{]}, percentage change on the baseline close to variable
transformation step

@@

The purpose of normalising is for cross-comparison. This step can get
criticism from analysts for \ldots{}

specifically for converting from a fitted distribution to normal score
via reverse CDF function, non-parametric formula, or empirical
approximation, a common step in many index: SPI, SSI, Z score. The
purpose of normalising is to convert the index into a standardised
series after all the steps for the ease of comparison.

Normalising is usually the last step

Restrict it to single variable, square root, log etc could be linearly,
also non-linear

change the shape of the variable

GAM, can you do additive model pairwise/ three-way

sometimes called feature extraction in the machine learning community
With drought indexes, the extraction of meaningful variables from the
original data is usually supported by the water balance model, for
example, in SPEI, the step that create \(d\) out of precipitation and
potential evapotranspiration (PET) has theoretical backup from {[}see
paper.{]}

Also include weighting

@@

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{amsmath}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Demo arXiv template},
  pdfauthor={H. Sherry Zhang; Collaborators},
  pdfkeywords={indexes, data pipeline, software design},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Demo arXiv template}
\author{H. Sherry Zhang \and Collaborators}
\date{}

\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

\emph{constructing time series index should also be encapsulated in the
framework}

Figure~\ref{fig-pipeline-steps} presents the index construction pipeline
from the raw data to the final indexes. The available intermediate steps
include temporal processing, spatial aggregation, variable
transformation, scaling, dimension reduction, normalising, benchmarking,
and simplification. These steps do not have to be arranged in sequence
and not all the steps are required to present in every index
construction. Spatial aggregation and temporal processing are operations
of the data in space and time Variable transformation, scaling, and
dimension reduction concerns operations on the multivariate aspect of
the data. Normalisation, benchmarking, and simplification can be viewed
as the ``post-processing'' of indexes, which addresses the
cross-comparison across different indexes and time, and index
communication.

More steps are involved before the mathematical construction of indexes.
This includes 1) defining a concept of interest, being useful for
decision making and 2) finding the relevant and measurable variables to
characterise the concept. These choices define the scope of the indexes
and the data availability for constructing an index but not the
construction itself, hence not included in the pipeline framework. The
framework proposed in this paper assumes the objective of an index has
been defined and relevant variables have been collected and quality
controlled.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\FunctionTok{include\_graphics}\NormalTok{(here}\SpecialCharTok{::}\FunctionTok{here}\NormalTok{(}\StringTok{"figures/pipeline{-}steps.png"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[H]

@@

\caption{\label{fig-pipeline-steps}Diagram of pipeline steps for index
construction.}

@@

\emph{Discussion: whether we should allow raw data to have different
spatial/ temporal resolution: This would require the notation
\(\mathbf{s}\) and \(\mathbf{t}\) to depend on \(p\)}

\emph{Initially my thought was to raw data only after aligning the
resolution: Another section on original data directly downloaded, can
have different spatial resolution, temporal granularity, data quality
problem. After processing them and align them together they become the
``raw data''}

Once variables are collected, they can have different spatial resolution
and temporal granularity. Let \(\mathbf{x}(\mathbf{s};\mathbf{t})\)
denote the raw data that contain the spatial, temporal, and multivariate
aspect: the spatial component
\(\mathbf{s} = (s_1, s_2, \cdots, s_n)^\prime\) is defined in the 2D
space: \(\mathbf{s} \in \mathcal{D}_s \subseteq \mathbb{R}^2\), the
temporal component \(\mathbf{t} = (t_1, t_2, \cdots, t_J)^\prime\) is
defined in the 1D space:
\(\mathbf{t} \in \mathcal{D}_t \subseteq \mathbb{R}\). Expanding on the
multivariate aspect, the data can be written as
\(\mathbf{x}(\mathbf{s}; \mathbf{t}) = (x_1(\mathbf{s}; \mathbf{t}), x_2(\mathbf{s}; \mathbf{t}), \cdots, x_P(\mathbf{s}; \mathbf{t}))^\prime\)

Temporal processing concerns the operations on the temporal aspect of
the data. It can be represented by
\(f_{\mathcal{\psi}}(x(\mathbf{s};\mathbf{t}))\), where
\(\psi \in \Psi \subseteq \mathbb{R}^{d_{\psi}}\) is the parameters
associated with the temporal operation and \(d_{\psi}\) is the number of
parameter of \(\psi\).

\begin{itemize}
\tightlist
\item
  purpose is to aggregate across time, smooth?
\end{itemize}

An example of temporal processing is to aggregate the data,
\(x(s_i; t_j)\), into \(k\)-month sum. This can be used to reflect how
the accumulated values change across time. In the drought index SPI,
precipitation can be aggregated at different scales to reflect
short-term, mid-term, and long-term drought severity. With \(\psi = k\),
the aggregated series can be written as
\(x(s_i;t_{j^\prime}) = \sum_{j = j^\prime-k+1}^{j^\prime}x(s_i; t_j)\)
where \(j^\prime\) is the time index after the aggregation.

@@

Sometimes, it can be beneficial to borrow information from nearby
spatial locations or there is the need to combine raster data from
satellite images with vector ground measures. Spatial aggregation can be
written as \(g_{\mathcal{\theta}}(x(\mathbf{s};\mathbf{t}))\), where
\(\theta \in \Theta \subseteq \mathbb{R}^{d_{\theta}}\) is the spatial
operation parameters and \(d_{\theta}\) is the number of parameter of
\(\theta\).

\begin{itemize}
\item
  \(\theta\) can be a constant or a derived from the data
\item
  3 by 3 aggregation, another purpose: the spatial grid is too fine,
  aggregate to higher level.
\end{itemize}

A variable transformation changes the shape of the variable. It can be
written as \(h_{\tau}(x(\mathbf{s};\mathbf{t}))\), where
\(\tau \in T \subseteq \mathbb{R}^{d_{\tau}}\) is the potential
parameter in the transformation, if any, and \(d_{\tau}\) is the number
of parameter of \(\tau\). Variable transformation involve square root
transformation, log transformation, and also include those non-linear
transformation.

@@

Scaling is a specific variable transformation on the scale of the data.
The purpose to treat scaling as a separate step from variable
transformation is to make the step explicit. Scaling is usually in the
form of \([x(s_i;t_j) - \alpha]/\gamma\). For example,
\(\alpha = \bar{x}(s; t)\) and \(\gamma = \sigma(s; t)\) is used in a
z-score standardisation to standardise the data by the its mean and
standard deviation across all time and all space units and
\(\alpha = \min[x(s_i, t_j)]\) and
\(\gamma = \max[x(s_i, t_j)] - \min[x(s_i, t_j)]\) is used to
standardise the variable into a unit interval.

When multiple variables are collected in space and time to construct
indices, they are often summarised with dimension reduction. Dimension
reduction
\(x_{p^*}(\mathbf{s}; \mathbf{t}) \rightarrow x_p(\mathbf{s}; \mathbf{t})\),
where \(p^* = 1, 2, \cdots, P^*\), \(p = 1, 2, \cdots, P\), and
\(P^* < P\). The most commonly used dimension reduction
techniquePprincipal Component Analysis (PCA), also called Empirical
Orthogonal Function (EOF), can be written as
\(x_{p^*}(\mathbf{s}; \mathbf{t}) = \sum_{p = 1}^{P}\lambda_{p}x_p(\mathbf{s};\mathbf{t})\)
where \(\lambda_p\) is the loading of the PC1, derived from maximising
the variance of the data given the constraint
\(\sum_{p=1}^P\lambda_p^2 = 1\).

@@

The purpose of normalising is for cross-comparison using the normal
score. This step can get criticism from analysts for \ldots{}

This step is usually in the end of the pipeline when the data have been
processed into a univariate series. When the data are not distributed
normally, a distributed is usually fitted before converting to the
normal score via the normal reverse CDF function: \(\Phi^{-1}(.)\):
\(\Phi^{-1}[F_{\eta}(x(\mathbf{s}; \mathbf{t}))]\), where
\(\eta \in H \subseteq \mathbb{R}^{d_{\eta}}\) is the distribution
parameter and \(d_{\eta}\) is the number of parameter of \(\eta\).
Sometimes, to account for the monthly variation in the data, the fitting
is done separately for each month and the noramlising step can be
written as \(\Phi^{-1}[F_{\eta}^m(x(\mathbf{s};t_{j^*}))]\), where
\(j^*\) is all the indexes that satisfy \(j^* \mod 12 = m\) for each
\(m = 0, 1,\cdots, 11\).

\begin{itemize}
\tightlist
\item
  In some cases, the distibution and inverse normal can be multivariate,
  i.e.~use copula.
\end{itemize}

@@

Benchmarking sets a constant value to allow the constructed index to be
compared across time. Here we denote it with \(u[x(s_i, t_j)]\) where
\(u\) is a scalar of interest in the index constructed, could be a
constant or a function of the data, i.e.~mean.

@@

In public communication, indexes are often delivered in categorical
grades, along with its underlying numerical values. The simplification
step prescribes how the continuous index values are converted into the
discrete grades. This process can be written with the piece-wise
function:

\[
\begin{cases}
C_0 & c_1 \leq (s_i; t_j) < c_0 \\
C_1 & c_2 \leq x(s_i; t_j) < c_1 \\
C_2 & c_3 \leq x(s_i; t_j) < c_2 \\
\cdots \\
C_z & c_z \leq x(s_i; t_j)
\end{cases}
\]

where \(C_0, C_1,\cdots ,C_z\) are the categories and
\(c_0, c_1, \cdots, c_z\) are the threshold value in each category. In
SPI, drought are sorted into four categories: mild drought:
\([-0.99, 0]\); moderate drought: \([-1.49, -1]\); severe drought:
\([-1.99, -1.5]\), and extreme drought: \([-\infty, -2]\). In this case,
\(C_0, C_1, C_2, C_3\) are the drought categories: mild, moderate,
severe, and extreme drought (\(z = 3\)) and
\(c_0 =0, c_1 = -1, c_2 = -1.5, c_3 = -2\) are the cutoff value for each
class.
